Paper template for {ICPhS} 2023 Prague

### 2024-04-27 manuscript/main.Rmd

Write something here

@@

There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules **(Institut d'Estudis Catalans, 2022)**, the [z] should be produced, but [s] is produced instead. This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such asthe so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p. 268)**. The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia. Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia **(Davidson, 2020)**.

@@

\subsection{The present study}
**INTRO**. Accordingly, the research questions and hypotheses are as follows:
1. What covert attitudes are linked to the perception Catalan intervocalic /s/ as either [s] or [z]?
2. What is the relationship between  individual social factors (mother tongue, province of origin) and covert attitudes towards Catalan intervocalic /s/?

@@

Participant responses to the statements *Parla bé català* 'S/he speaks proper Catalan' and *És d'origen català* 'S/he is of Catalan origin' showed similar results, as illustrated in Figures \ref{fig:pbc} and {fig:oc}. Those who have both Catalan and Spanish as their L1 rated the speaker more positively regardless of phoneme, compared to those whose L1 is either Catalan or Spanish.

@@

\section{Discussion and Conclusions}
The aim of this exploration study was to investigate whether phoneme production ([s] vs [z]), mother tongue (Catalan vs Spanish vs both), and province of origin (Girona vs Barcelona) influenced the attitudes of the participants towards a Catalan native speaker. The results suggest that these predictors have a different effect on the ratings of the speaker depending on the social characteristic under evaluation.

Contrary to what we expected, phoneme was not a significant predictor. One explanation for this divergent pattern could be that either participants could not distinguish between [s] and [z] or, if they could, this difference in production was not changing their attitudes towards the speaker. This finding is interesting considering that, in our stimuli, the production of the voiceless alveolar fricative in intervocalic position has been considered a result of Catalan in contact with Spanish **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p. 268)**, and, from a prescriptivist point of view, it should be avoided **(Institut d'Estudis Catalans, 2022)**. Despite the lack of significance of this predictor, it should be mentioned that the speaker received more positive ratings when the [z] was being produced. Although at this point we can only hypothesize, it could be that the combination of the phoneme /s/ with the precedent vowel, might give the listener more cues and have a stronger social meaning. Future studies should address this issue in order to explore if the combination of vowel + /s/ carries other social connotations. **REVISAR**

**CONTINUAR**

>>>>

Catalan and Spanish are two languages that have coexisted for centuries, mainly in eastern Spain, along the Mediterranean cost **(Hualde, 1992)**. It is safe to assume that there is a bidirectional influence between these languages: in that region, Catalan has influenced Spanish and Spanish has influenced Catalan in all linguistic levels **(Davidson, 2014, 2015)**. A crucial area of cross-linguistic influence might be the production of sounds, especially in those sounds who have different distribution in Catalan and Spanish. For example, previous studies have found that, in Barcelona, the production of Catalan vowels is different based on the L1 of the speaker, being it less target-like in L1 Spanish speakers compared to L1 Catalan speakers **(CITAR SEBASTIAN GALLES, etc)**.

As for how Catalan people perceive these differences in production and, thus, their attitudes towards a less Catalan-like production of a given phoneme little is known. The aim of this study is to examine the covert attitudes towards the production of intervocalic /s/, as either [s] (disfavored by prescriptivist norms) or [z] (preferred by prescriptivist norms) by L1 Catalan, L1 Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched guise test **(Lambert, 1967)**. The alveolar fricative was chosen because it exists in both Catalan and Spanish inventories, but their contextual distribution is different **(Llisterri, 2023)**.

**PONER ALGO MÁS**

@@

There is some evidence of the influence of Spanish on the production of the Catalan alveolar fricative, in contexts where, according to prescriptivist rules **(Institut d'Estudis Catalans, 2022)**, the [z] should be produced, but [s] is produced instead. This phenomenon has mainly been attested in those areas where there has been an extended contact between Catalan and Spanish, such as the so-called "franja de Aragón" (borderland between Catalonia and Aragon), some areas of the Valencian Community, and Barcelona and other urban areas **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p. 268)**. The opposite pattern has also been observed - that is, the voicing of the alveolar fricative in the Spanish spoken in Barcelona and Valencia. Interestingly, although in both Barcelona and Valencia Catalan and Spanish coexist and influence each other, it seems that Catalan has a greater influence on Spanish in Barcelona, but  Spanish has a greater influence on Catalan in Valencia **(Davidson, 2020)**.

@@

\subsection{The present study}
**INTRO**.

Accordingly, the research questions and hypotheses are as follows:
1. What covert attitudes are linked to the perception Catalan intervocalic /s/ as either [s] or [z]?
2. What is the relationship between  individual social factors (mother tongue, province of origin) and covert attitudes towards Catalan intervocalic /s/?

@@

Participant responses to the statements *Parla bé català* 'S/he speaks proper Catalan' and *És d'origen català* 'S/he is of Catalan origin' show similar results, as illustrated in Figures \ref{fig:pbc} and {fig:oc}. Those who have both Catalan and Spanish as their L1 rate the speaker more positively regardless of phoneme, compared to those whose L1 is either Catalan or Spanish. In addition, although it is not statistically significant, participants give higher ratings to the speaker when they are listening to those statements with [z].

@@

\section{Discussion and Conclusions}
The aim of this exploration study was to investigate whether phoneme production ([s] vs [z]), mother tongue (Catalan vs Spanish vs both), and province of origin (Girona vs Barcelona) influenced the attitudes of the participants towards a native speaker of Central Catalan. The results suggest that these predictors have a different effect on the ratings of the speaker depending on the social characteristic under evaluation.

Contrary to what we expected, phoneme was not a significant predictor. One explanation for this divergent pattern could be that either participants could not distinguish between [s] and [z] **(Davidson, 2019)** or, if they could, this difference in production was not changing their attitudes towards the speaker **(Sinner, 2002)**. This finding is interesting considering that the production of the voiceless alveolar fricative in intervocalic position has been considered a result of Catalan in contact with Spanish **(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996, p. 268)**, and, from a prescriptivist point of view, it should be avoided **(Institut d'Estudis Catalans, 2022)**. Despite the lack of significance of this predictor, it should be mentioned that the speaker received more positive ratings when the [z] was being produced. The fact that phoneme was not an important factor contrasts with **Davidson (2020)**, who analyzed covert and overt attitudes towards Catalonian Spanish and found that [z] was a significant predictor: it was considered a sociolinguistic indicator of Catalan-Spanish bilingual membership. Although at this point we can only hypothesize, it could be that [z] is perceptually more salient in Spanish than in Catalan, as the Spanish inventory does not have the phoneme /z/, but voicing of /s/ might partially occur **(Benet et al., 2012; Hualde & Colina, 2015; Hualde et al., 2020)**. Another explanation could be that the /s/ alone does not provide enough social information; rather, the combination of the phoneme /s/ with the precedent vowel might give the listener more cues and have a stronger social meaning. Future studies should address this issue in order to explore if the combination of vowel + /s/ carries other social connotations than just the /s/ alone.

We expected mother tongue to be a strong predictor of the attitudes towards the production of intervocalic /s/, specifically, that L1 Catalan and bilingual speakers would give higher rates to those statements with [z]. This hypothesis was only partially confirmed. As reported above, phoneme was not a factor influencing the attitudes, and mother tongue was only a significant predictor in the two statements related to Catalan identity, namely *Parla bé català* 'S/he speaks proper Catalan' and *És d'origen català* 'S/he is of Catalan origin'. Interestingly, bilinguals gave higher ratings to the speaker, followed by L1 Catalan participants and L1 Spanish participants. A potential explanation for this finding, that should be verified in future research, could be that L1 Catalan speakers have specific ideas (higher standards) towards the 'proper' Catalan. In addition, some studies have suggested that those Catalan people who grow up in Spanish-speaking households typically exhibit more negative attitudes towards Catalan **(CITATIONS)**, which is consistent with our results.

Lastly, province was also found as an important predictor, but only for some statements. In two of the solidarity statements (*És simpàtic* 'S/he is nice'; *És de fiar* 'S/he is trustworthy'), the speaker was rated more positively by those participants from the province of Barcelona. This finding might suggest that they identify with the speaker; thus, giving her higher ratings **BUSCAR CITA**. Interestingly, the speaker that created the stimuli is from the province of Girona, but has spent the last 11 years of her life living outside that area, either in Barcelona or in another country. We asked her to read the sentences using a Central Catalan 'neutral accent', which people might have potentially related with Barcelona. This 'neutral accent' does not show any of the prototypical characteristics of the *accent gironí*, a subvariety within Central Catalan spoken in some areas of Girona **(Gros, 2024)**.

In summary, our findings suggest that the relationship between covert attitudes and speaker ratings is not straightforward, as the speaker was evaluated differently based on the social characteristic, with varying factors influencing such ratings (mother tongue and province). Crucially, phoneme production did not had an effect on the ratings, even though prescriptivist norms disfavor the realization of [s] in intervocalic position. Future studies should include a larger participant pool, as well as participants from other regions in Catalonia, such as Lleida and Tarragona. In addition, examining other phoneme productions that emerge through the contact of Catalan with Spanish could give us a better understanding of the factors that affect the attitudes towards the Catalan language, and potential sociolinguistic stereotypes and indicators related to Catalan identity.
